ExpressionView - an interactive viewer for biclusters identified in gene expression data

### 2010-02-04 isanote/answers_to_reviewers.tex

\documentclass[a4paper]{article}
\usepackage{url}

\title{Modular analysis of gene expression data with R\\
  Answers to reviewers}
\author{G\'abor Cs\'ardi, Zolt\'an Kutalik and Sven Bergmann}

\begin{document}

\maketitle

\section{Reviewer 1}

\begin{quote}
Comments to the Author

This Application Note presents a software package that implements the
Iterative Signature Algorithm and provides a convenient visualization
of it’s results. The method is useful and the software may be of help
to interested users. 
\end{quote}

\begin{quote}
I have two problems with the Note in its present form.
The first may be due to my misunderstanding of the method. My
understanding was that some threshold Tg is applied on the genes and
another one, Tc, on the conditions. The aim is to reach a bicluster (a
set of conditions C and a set of genes G) such that G contains only
genes with expression above Tg for all the conditions of C, and
conditions are included in C only if the expression of all genes of G
exceeds Tc (one can replace a lower threshold with an upper one). The
heatmap of the Figure implies that my understanding was incorrect –
the conditions satisfy a double threshold (expressions are either
above or below some thresholds) and the genes have expressions above a
Tg in some conditions or below some other Tg’ in other conditions. 
I urge the authors to define what specifies or defines a bicluster –
do it briefly and sacrifice a few lines from elsewhere. 
\end{quote}

\begin{quote}
My second concern is minor – the abstract implies that the ISA is
special in allowing overlapping cluster assignments. This is incorrect
– there are other biclustering methods that do this. If I am not
mistaken, all three original biclustering methods that came out in
2000 (Cheng, Getz and Califano) allow this. 
\end{quote}

\section{Reviewer 2}

\begin{quote}
Comments to the Author

Summary

This manuscript describes two software packages for GNU R, for
discovering biclusters (here called transcription modules. see [1] for
a survey) in a given gene expression matrix based on the Iterative
Signature Algorithm (ISA) [2]. The first package, called isa2,
contains the implementation of the basic ISA, and can be used to
analyze any tabular data. The second package, called eisa, is built on
isa2. It adds support to standard BioConductor [3] data structures,
and contains gene expression specific tools for visualization of
biclusters. Both packages support execution of ISA with default
parameters, and also execution of ISA step by step with user-specified
parameters. The steps of ISA are normalization, iteration, filtering
of modules and enrichment calculation. In the current note, the
authors present both packages, and describe how to install and run
them step by step, using the acute lymphoblastic leukemia data set
(ALL) [4] as an example. 
\end{quote}

\begin{quote}
Overall evaluation

The manuscript and its accompanying tutorials are well written and
give an adequate theoretical background to ISA. The tutorials include
many detailed and visualized examples that make learning the package
quite easy for an R user. However, the following major points should
be taken into consideration: 
\end{quote}

\begin{quote}
1.      The manuscript and the website lack explanations about how the
authors implemented ISA and how they used the advantages of R. The
authors emphasize that the packages are highly optimized and run
extremely fast even for very large datasets.  Therefore, the
accompanying material should contain a summary describing how this
efficiency was achieved. In addition, a graph showing execution time
as a function of dataset size should be given. 
\end{quote}

\begin{quote}
2.      The tutorials give examples of execution on a single rather
old dataset (ALL) which is quite small in comparison to datasets of
recent years. Here the authors should give results for ISA execution
on more up-to-date datasets, and on heterogeneous datasets containing
profiles from multiple studies. 
\end{quote}

\begin{quote}
To summarize, the software is well motivated and enables users of the
R environment to utilize ISA. I suggest accepting the manuscript
conditional on solving the issues mentioned in this report. 
\end{quote}

\begin{quote}
Additional important remarks:

1.      The manuscript should contain references to additional
biclustering implementations, in R or in other environments, so that
the user could choose the one that best fits his needs in terms of
input/output, species, etc. Examples are SAMBA (via the Expander
software) [5,6] and BicAt [7]. 
\end{quote}

\begin{quote}
2.      The sample score plots example given in the EISA tutorial
(\url{http://www2.unil.ch/cbg/index.php?title=EISA\_tutorial}, section 6.5)
is not sufficiently clear. What are the scores represented by the
horizontal lines and how can one calculate them? Moreover, how can one
separate B-cell and T-cell leukemia samples using this plot? 
\end{quote}

\begin{quote}
3.      It is not quite clear what is the difference between the ISA
implementation included in the isa2 package and the ISA implementation
included in the eisa package. 
\end{quote}

\begin{quote}
4.      It should be clarified who are the potential users for the ISA
implementation for R environment. Assuming these are mainly
biologists, bioinformaticians and biostatisticians (and not only
programmers), more technical details of how to work with the packages
should be given. The following points are important examples: 
a.      Since the ISA implementation in eisa package imports data from
another package (e.g. hgu95av2.db), a tutorial for how to create such
a .db package from a given expression data text file should be given
(or referred to). 
\end{quote}

\begin{quote}
b.      Since both ISA packages use various data structures, a short
explanation of how to construct and use them should be given (or
references to such explanation), and also how to export data from
these structures to external files. 
\end{quote}

\begin{quote}
Minor remarks:

1.      The authors describe methods that perform enrichment tests for
the gene sets corresponding to the ISA modules against various
databases. In particular, they perform enrichment analysis against GO
database, KEGG pathway database, chromosomes and the TargetScan [8]
database. One may also want to investigate the correspondence of
biclusters and protein-protein interaction networks (which have been
derived from other types of data than gene expression data), e.g. the
DIP database [9]. An example for such analysis can be found in [10]. 
\end{quote}

\begin{quote}
2.      The heatmap image of a transcription module obtained by any
one of the two methods presented in the tutorial should include a
color legend. 
\end{quote}

\begin{quote}
3.      Since ISA function accepts an ExpressionSet, one must obtain
annotations for its expression data file in order to run ISA function
on it. Annotations should not be a prerequisite for running ISA. 
\end{quote}

\begin{quote}
4.      In section 10.8 it is mentioned that two significant modules
were found using random seeds and two significant (separating) modules
were found using smart seeds. However, in figure 7 it is noted that
only one module found using smart seeds. At the end of the section it
is mistakenly written "As it turns out, one modules found". 
\end{quote}

\begin{quote}
References:
[1] A. Tanay, R. Sharan, and R. Shamir. Biclustering algorithms: A
survey. In E. by Srinivas Aluru, editor, In Handbook of Computational
Molecular Biology. Chapman \& Hall/CRC, Computer and Information
Science Series, 2005.

[2] Bergmann, S., Ihmels, J., and Barkai, N. (2003). Iterative
signature algorithm for the analysis of large-scale gene expression
data. Phys Rev E Nonlin Soft Matter Phys, page 031902.

[3] Gentleman, R. C., Carey, V. J., Bates, D. M., et
al. (2004). Bioconductor: Open software development for computational
biology and bioinformatics. Genome Biology, 5, R80.

[4] Chiaretti, S., Li, X., Gentleman, R., Vitale, A., Vignetti, M.,
Mandelli, F., Ritz, J., and Foa, R. (2004). Gene expression profile of
adult t-cell acute lymphocytic leukemia identifies distinct subsets of
patients with different response to therapy and survival. Blood,
103(7).

[5] R. Sharan, A. Maron-Katz, N. Arbili, and R. Shamir.EXPANDER:
EXPression ANalyzer and DisplayER, 2002.Software package, Tel-Aviv
University, \url{http://www.cs.tau.ac.il/~rshamir/expander/expander.html}.

[6] A. Tanay, R. Sharan, M. Kupiec, and R. Shamir. Revealing
modularity and organization in the yeast molecular network by
integrated analysis of highly heterogeneous genomewide data. Proc Natl
Acad Sci U S A., 101(9):2981–6, 2004.

[7] Barkow, S., et al. (2006) BicAT: a biclustering analysis
toolbox. Bioinformatics, 22, 1282–1283

[8] Targetscan [\url{http://www.targetscan.org/}]

[9] Salwinski, L., et al. (2004) The Database of Interacting Proteins:
2004 update. Nucleic Acids Res, . 32, D449–D451

[10] Prelic,A. et al. (2006) A systematic comparison and evaluation of
biclustering methods for gene expression data. Bioinformatics, 22,
1122–1129. 
\end{quote}

\end{document}

### 2010-03-14 expressionviewnote/expressionview.tex
(R)ExpressionView: BioC modifications.

\documentclass[round]{bioinfo}

@@

\usepackage{amstext}
\usepackage{url}
% \usepackage[colorlinks,citecolor=blue,linkcolor=red,urlcolor=blue]{hyperref}
\newcommand{\thefirstpage}{1} \newcommand{\thelastpage}{1}

\copyrightyear{2010}
\pubyear{2010}

@@

\firstpage{1}
\application{}
\title[ExpressionView]{ExpressionView - an interactive viewer for
  modules identified in gene expression data} 
\author[Andreas L\"uscher, G\'abor Cs\'ardi, Aitana Morton de
Lachapelle, Zolt\'an Kutalik, and Sven Bergmann]{Andreas
  L\"uscher$^{1\dagger}$, G\'abor Cs\'ardi$^{1,2\dagger}$, Aitana Morton de
  Lachapelle$^{1,2\dagger}$, Zolt\'an Kutalik$^{1,2\dagger}$, 
  Bastian Peter$^{1,2}$
  and Sven Bergmann$^{1,2}$}
\address{
  $^{1}$Swiss Institute of Bioinformatics, Lausanne, Switzerland\\
  $^{2}$Department of Medical Genetics, University of Lausanne,
  Lausanne, Switzerland\\
  +: equal contribution
}

@@

\begin{abstract}
\section{Summary:}
ExpressionView is an R package that provides an interactive graphical
environment to explore transcription modules identified in gene expression
data. A sophisticated ordering algorithm is used to present the
modules with the expression in a visually appealing layout that provides an intuitive
summary of the results. From this overview, the user can select
individual modules and access all the biologically relevant metadata
associated with them.

\section{Availability:} 
%\href{http://www.unil.ch/cbg/ExpressionView}{R 
%  package with open source license,
\url{http://www.unil.ch/cbg/ExpressionView}
%}

\section{Contact:} \href{Sven.Bergmann@unil.ch}{Sven.Bergmann@unil.ch}

\section{Supplementary information:}
Screenshots, tutorials and sample data sets can be found on the
ExpressionView \href{http://www.unil.ch/cbg/ExpressionView}{website}. 
\end{abstract}

\vspace*{-7pt}
\section{Introduction}
Biclustering is an unsupervised data analysis method which is
frequently used to explore microarray data. Biclustering algorithms
identify co-expressed groups of
genes and conditions (samples) where this co-expression occurs from
collections of expression profiles. We refer to such groups as modules. While 
there is a multitude of biclustering software available~(see in
\cite{madeira04,ihmels04}), packages with intuitive interfaces that
allow for an interactive exploration of the results are
sparse~\citep{santamaria08}. ExpressionView is designed to close this
gap, and facilitate the collaboration between bioinformaticians and
life scientists without consolidated programming
experience. Implemented as an R package, the user can apply all the
powerful microarray analysis tools provided by the
Bioconductor~\citep{gentleman04} project before exporting the data to
a platform independent Adobe Flash applet. This applet
provides an integrated, intuitive and interactive visualization of
the modules together with the underlying gene expression data and
additional information on the genes and conditions.

\vspace*{-7pt}
\section{Package design and workflow}
The ExpressionView package contains two independent parts: An ordering
algorithm implemented in C++ and an Adobe Flash applet written in
ActionScript and Adobe Flex. Fig.~\ref{fig:workflow} schematically
summarises the typical workflow that uses the gene expression data
available as a Bioconductor ExpressionSet and combines it with the
biclustering results to produce an ExpressionView data file that can
be explored with the Flash applet. In the following, we describe the
different steps in more detail. 
\begin{figure*}[!tpb]
\centerline{\includegraphics[width=0.7\linewidth]{fig1-crop}}
\caption{How to use ExpressionView: Starting from gene expression data
  in the form of a Bioconductor ExpressionSet, the user first runs a
  biclustering algorithm to determine co-expressed groups of genes and
  the relevant samples. In a second step, the rows and columns of the gene
  expression matrix are rearranged to produce an easily readable
  overview of the results. The last step consists of combining the
  gene expression data and its associated metadata with the results
  from the biclustering and produce an ExpressionView data file that
  can be explored with the interactive Flash
  applet.}\label{fig:workflow} 
\end{figure*}
For users who prefer to install the viewer as a standalone program, we
also provide an Adobe AIR version that can be downloaded from the
\href{http://www.unil.ch/cbg/ExpressionView}{website}.

\subsection{Gene expression data and modules}
ExpressionView is designed to work with gene expression data in the
form of a Bioconductor ExpressionSet. This class provides a
user-friendly way to access the actual gene expression matrix and its
associated metadata. There is a variety of biclustering algorithms
described in the literature~\citep{madeira04,prelic06}, several of
which have been implemented as R packages. ExpressionView can treat
results obtained by the Iterative Signature
Algorithm~\citep{bergmann03,csardi09} and any of the methods available
in the Biclust package~\citep{kaiser08}. Since the structure of
biclustering results is independent of the algorithm, an extension to
other methods is straightforward.

\subsection{Reordering genes and conditions}
To present the collection of possibly overlapping modules in a visually
appealing form, it is necessary to reorder the rows (genes) and
columns (conditions) of the
gene expression matrix in such a way that biclusters form contiguous
rectangles. Since it is in general impossible to find such an
arrangement for more than two mutually overlapping modules, we propose
here an approximate solution that optimizes the arrangement within the
original data by maximising the total area of the largest contiguous
module subsets. (An alternative would be to repeat rows and columns as
necessary~\citep{grothaus06}, but for many modules this will results in
a very large expression matrix.)

This optimization task is an interesting problem on its own, which to
the best of our knowledge has not been studied in the literature. Thus
we briefly outline our strategy here (see our website for details). 
The reordering of the rows and the columns do not depend on each
other, so the same optimization method can be applied sequentially,
say first to rows and then to columns. For a given order of the
elements (either genes or conditions) we compute for each module the
size of the largest contiguous sequence of elements (i.e. the maximal
number of neighboring elements $N^\text{max}_i$). Then as a measure of the
quality $Q$ of the order we sum this quantity over all modules
($Q=\sum_i N^\text{max}_i$). The optimization procedure for $Q$ starts from
some initial order. Two operations are then applied to this order: (1)
Permutations that exchange two elements within a module and (2) shifts
of a sequence of elements within a module to a different position. We
use a greedy iterative scheme that performs these operations at
randomly selected positions and keeps the new order if it improves
$Q$. The algorithm stops if after a given number of operations no
significant improvement of $Q$ is achieved.

To determine the efficiency of this method, we have studied a large
number of perfectly orderable, but initially scrambled, test cases. We
find that in almost every situation, the proposed algorithm finds an
order that recovers more than 99\% of the score of the optimal
solution and in most cases, it recovers the correct
alignment. For random samples, which are more representative for
actual gene expression data, the execution time increases polynomially
with the number of clusters $m$ as ${\mathcal O}(m^\alpha)$, where
$\alpha \in [1.6, 2]$, almost independent of the number of elements
$n$. For a given number of clusters, we find ${\mathcal O}(n^\alpha)$,
with $\alpha \in [2.5, 2.7]$.

\subsection{Export}
Once the optimal order is determined, the program rearranges the gene
expression matrix accordingly and exports all the relevant information
into an XML file. This format has the advantage of being
self-explanatory and extensible. In addition, the files can be read
and edited with any text editor. The actual gene expression data is
rounded to two digit precision and encoded in Base64. The structure of
the XML file is described in more detail on the ExpressionView
\href{http://www.unil.ch/cbg/ExpressionView}{website}. In most cases,
these files are of the order of a few Megabytes and can thus easily be
sent by e-mail.

\subsection{Visualisation}
The core application of this package is the Flash applet that allows
the user to interactively explore the gene expression data and the
modules. The Flash architecture has the advantage of
producing a platform and browser independent applet that can be
launched from any computer with even the most restrictive user
rights. A screenshot is shown in Fig.~\ref{fig:workflow}. The
interface is divided in two parts: On the left-hand side, the user
finds the gene expression data in the common heat map form, on top of
which the modules are overlaid. On the right-hand side, the metadata
associated with the gene expression data and the results of the
enrichment calculations, GO~\citep{ashburner00} and
KEGG~\citep{kanehisa04}, are shown. Wherever possible, these elements
are linked to the corresponding databases. The interface essentially
behaves as an image viewer, allowing the user to zoom and pan around
the expression data, getting instant feedback on the selected
item. The currently selected region can be exported to a PDF file,
helpful to prepare figures for presentations or articles, or a simple
text file.

\vspace*{-7pt}
\section*{Acknowledgement}

\paragraph{Funding\textcolon} The authors are grateful to the Swiss
Institute of Bioinformatics, the Swiss National Science Foundation
(3100AO-116323/1) and the European Framework Project 6 (through
the EuroDia and AnEuploidy projects).

@@

\vspace*{-7pt}
\bibliographystyle{natbib}
\bibliography{expressionview}

>>>>

\documentclass{bioinfo}

@@

\usepackage[dvipdfm,colorlinks,bookmarks=false,citecolor=blue,linkcolor=red,urlcolor=blue]{hyperref}
\newcommand{\thefirstpage}{1} \newcommand{\thelastpage}{1}

\copyrightyear{2009}
\pubyear{2009}

@@

\firstpage{1}
\application{}
\title[ExpressionView]{ExpressionView - an interactive viewer for biclusters identified in gene expression data}
\author[Andreas L\"uscher, G\'abor Cs\'ardi, Aitana Morton de Lachapelle, Zolt\'an Kutalik, and Sven Bergmann]{Andreas L\"uscher$^1$, G\'abor Cs\'ardi$^{1,2}$, Aitana Morton de Lachapelle$^{1,2}$, Zolt\'an Kutalik$^{1,2}$, and Sven Bergmann$^{1,2}$\footnote{to whom correspondence should be addressed}}
\address{
	$^{1}$Swiss Institute of Bioinformatics, Lausanne, Switzerland\\
	$^{2}$Department of Medical Genetics, University of Lausanne, Lausanne, Switzerland
}

@@

\begin{abstract}

\section{Summary:}
ExpressionView is an R package that provides an interactive environment to explore biclusters identified in gene expression data. A sophisticated ordering algorithm is used to present the biclusters in a visually appealing layout that provides an intuitive summary of the results. From this overview, the user can select individual biclusters and access all the biologically relevant data associated with it. The package is aimed to facilitate the collaboration between bioinformaticians and life scientists who are not familiar with the R language.

\section{Availability:} 
\href{http://www2.unil.ch/cbg/index.php?title=ExpressionView}{R package with open source license}

\section{Contact:} 
\href{andreas.luescher@a3.epfl.ch}{Andreas L\"uscher}

\section{Supplementary information:}
Screenshots, tutorials and sample data sets can be found on the ExpressionView \href{http://www2.unil.ch/cbg/index.php?title=ExpressionView}{website}.

\end{abstract}

\section{Introduction}
Biclustering is an unsupervised learning technique used to analyse microarray data. It consists in identifying co-regulated groups of genes and conditions according to their expression profile. While there is a multitude of biclustering software available~\citep{madeira04}, packages with intuitive interfaces that allow the user to interactively explore the results are sparse~\citep{santamaria08}. ExpressionView is designed to close this gap, hereby facilitating the collaboration between bioinformaticians and life scientists without consolidated programming experience. Implemented as an R package, the user can apply all the powerful microarray analysis tools provided by the Bioconductor~\citep{gentleman04} project before exporting the data to a platform independent Adobe Flash applet that allows him and his co-workers to visualise the biclusters together with the underlying gene expression data.

\section{Package design and workflow}
The ExpressionView package contains two independent parts: An ordering algorithm implemented in C++ and a Adobe Flash applet written in ActionScript and Adobe Flex. Fig.~\ref{fig:workflow} schematically summarises the typical workflow that uses the genomic data available as a Bioconductor ExpressionSet and combines it with the biclustering results to produce an ExpressionView data file that can be explored with the Flash applet. In the following, we describe the different steps in more detail.
\begin{figure*}[!tpb]
\centerline{\includegraphics[width=0.9\linewidth]{fig1-crop}}
\caption{How to use ExpressionView: Starting from gene expression data in the form of a Bioconductor ExpressionSet, the user first runs a biclustering algorithm to determine co-regulated groups of genes and samples. In a second step, the rows and columns of the gene expression matrix are rearranged to produce an easily readable overview of the results. The last step consist in combining the gene expression data and its associated metadata with the results from the biclustering and produce an ExpressionView data file that can be explored with the interactive Flash applet.}\label{fig:workflow}
\end{figure*}
For users who prefer to install the viewer as a standalone program, we also provide a platform independent Adobe AIR version that can be downloaded from the \href{http://www2.unil.ch/cbg/index.php?title=ExpressionView}{website}.

\subsection{Gene expression data and biclusters}
ExpressionView is designed to work with gene expression data in the form of a Bioconductor ExpressionSet. This class provides a user-friendly way to access the actual gene expression matrix and its associated metadata. There is a variety of biclustering algorithms described in the literature~\citep{madeira04,prelic06}, several of which have been implemented as R packages. ExpressionView can treat results obtained by the iterative signature algorithm (ISA)~\citep{bergmann03,csardi09} and the methods available in the Biclust package~\citep{kaiser08}. Since the structure of biclustering results is independent of the algorithm, an extension to other methods is straightforward.

\subsection{Order}
To present the tens of possibly overlapping biclusters in a visually appealing form, it is necessary to reorder the rows and columns of the gene expression matrix in such a way that biclusters form contiguous rectangles. Since for more than two mutually overlapping biclusters, it is in general impossible to find such an arrangement, one has to make concessions. In contrast to \cite{grothaus06}, who proposed to repeat rows and columns as necessary to achieve this goal, we prefer to optimise the arrangement within the original data by maximising the area of the largest contiguous biclusters.

Since this reordering is an interesting problem on its own, which to the best of our knowledge has not been studied in the literature, we briefly outline our strategy here. Noting that the rows and the columns can be ordered separately, the problem reduces to finding the optimal arrangement of a set of $n$ elements that are part of at least one of $m$ clusters. The quality of the order is defined as the sum over the maximal number of neighbouring elements in every cluster and the tasks thus consists in maximizing this quantity by changing the order of the elements. Starting from an initial configuration determined by similarity, this is achieved by either shifting well aligned subsequences of a given cluster, herby enlarging the longest contiguous part, or permuting the individual elements of a given subcluster.

To determine the efficiency of this method, we have studied a large number of perfectly orderable, but initially scrambled, test cases. We find that in almost every situation, the proposed algorithm finds an order that recovers more than 99\% of the score of the optimal solution and in most cases, it actually recovers the correct alignment. For random samples, which are more representative for actual gene expression data than orderable situations, execution time increases polynomially with the number of clusters $m$ as ${\mathcal O}(m^\alpha)$, where $\alpha \in [1.6, 2]$, almost independent of the number of elements $n$. For a given number of clusters, we find ${\mathcal O}(n^\alpha)$, with $\alpha \in [2.5, 2.7]$.

\subsection{Export}
Once the optimal order is determined, the program rearranges the gene expression matrix accordingly and exports all the relevant information into an XML file. This format has the advantage of being self-explanatory and extensible. In addition, the files can be read and edited with any text editor. The actual gene expression data is rounded to two digit precision and encoded in Base64. The structure of the XML file is described in more detail on the ExpressionView \href{http://www2.unil.ch/cbg/index.php?title=ExpressionView}{website}. In most cases, these files are of the order of a few Megabytes and can thus easily be sent to co-workers.

\subsection{Visualisation}
The core application of this package is the Flash applet that allows the user to interactively explore the gene expression data and the biclusters. Relying on the Flash architecture has the advantage of producing a platform and browser independent applet that can be launched from any computer with even the most restrictive user rights. A screenshot is shown in Fig.~\ref{fig:workflow}. The interface is divided in two parts: On the left-hand side, the user finds the gene expression data in the common heat map form, on top of which the biclusters are overlaid. On the right-hand side, the metadata associated with the gene expression data and the results of the enrichment calculations [GO (Gene Ontology)~\citep{ashburner00} and KEGG (Kyoto Encyclopedia of Genes and Genomes)~\citep{kanehisa04}] are shown. Wherever possible, these elements are linked to the corresponding databases. The interface essentially behaves as an image viewer, allowing the user to zoom and pan around the expression data, getting instant feedback on the selected item. The currently selected region can be exported to a PDF file, helpful to prepare figures for presentations or articles, or a simple text file.\\

%\section*{Acknowledgement}

\paragraph{Funding\textcolon} This work was supported by ...

@@

\begin{thebibliography}{6}

\bibitem[Ashburner et~al.(2000)]{ashburner00}
Ashburner,M. \emph{et al.} (2000)
Gene Ontology: tool for the unification of biology.
\href{http://dx.doi.org/10.1038/75556}{\emph{Nature Genetics}, {\bf 25}, 25 -- 29.}

\bibitem[Bergmann et~al.(2003)]{bergmann03}
Bergmann,S., Ihmels,J. and Barkai,N. (2003)
Iterative signature algorithm for the analysis of large-scale gene expression data.
\href{http://dx.doi.org/10.1103/PhysRevE.67.031902 }{\emph{Physical Review E}, {\bf 67}, 031902.}

\bibitem[Cs{\'a}rdi et~al.(2009)]{csardi09}
Cs{\'a}rdi,G., Kutalik,Z. and Bergmann,S. (2009)
Modular analysis of gene expression data with R.
\href{http://dx.doi.org/}{\emph{Bioinformatics}, ...}

\bibitem[Gentleman et~al.(2004)]{gentleman04}
Gentleman,R.C., \emph{et al.} (2004)
Bioconductor: open software development for computational biology and bioinformatics.
\href{http://dx.doi.org/10.1186/gb-2004-5-10-r80}{\emph{Genome Biology}, {\bf 5}, R80.}

\bibitem[Grothaus et~al.(2006)]{grothaus06}
Grothaus,G., Mufti,A., and Murali,TM. (2006)
Automatic layout and visualization of biclusters.
\href{http://dx.doi.org/10.1186/1748-7188-1-15}{\emph{Algorithms for Molecular Biology}, {\bf 1}, 15.}

\bibitem[Kaiser and Leisch(2008)]{kaiser08}
Kaiser,S. and Leisch,F. (2008)
A toolbox for bicluster analysis in R.
\href{http://epub.ub.uni-muenchen.de/3293/}{Technical Report 028, Department of Statistics, University of Munich}.

\bibitem[Kanehisa et~al.(2004)]{kanehisa04}
Kanehisa,M. \emph{et al.} (2004)
The KEGG resource for deciphering the genome.
\href{http://dx.doi.org/10.1093/nar/gkh063}{\emph{Nucleic Acids Res.}, {\bf 32} (Database special issue), D277.}

\bibitem[Madeira and Oliveira(2004)]{madeira04}
Madeira,S.C. and Oliveira,A.L. (2004)
Biclustering algorithms for biological data analysis: A survey.
\href{http://dx.doi.org/10.1109/TCBB.2004.2}{\emph{IEEE/ACM Transactions on Computational Biology and Bioinformatics}, {\bf 1}, 24 -- 45.}

\bibitem[Prelic et~al.(2006)]{prelic06}
Prelic,A., \emph{et~al.} 
A systematic comparison and evaluation of biclustering methods for gene expression data 
\href{http://dx.doi.org/10.1093/bioinformatics/bt1060}{\emph{Bioinformatics}, {\bf 22}(9), 1122 -- 1129.}

\bibitem[Santamar\'ia et~al.(2008)]{santamaria08}
Santamar\'ia,R., Ther\'on,R. and Quintales,L. (2008)
BicOverlapper: A tool for bicluster visualization. 
\href{http://dx.doi.org/10.1093/bioinformatics/btn076}{\emph{Bioinformatics}, {\bf 24}, 1212 -- 1213.}

\end{thebibliography}
